---
title: "Assignment10"
author: "Morgann Plante"
date: "10/14/2020"
output: html_document
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment10.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

-------

1. Install the package `mlbench` and use the follows to import the data

```{r}
library(mlbench)
data(PimaIndiansDiabetes)
df <- PimaIndiansDiabetes
```

- Set seed to be 2020. 
- Partition the data into 80% training and 20% testing.  
```{r}
library(caret)
set.seed(2020)
splitIndex <- createDataPartition(df$diabetes , p = .80, 
                                  list = FALSE)
df_train <- df[ splitIndex,]
df_test <- df[-splitIndex,]
```

-------

2. Practice Decision Tree.  Do the follows:

  - Use `rpart` package, create a decision tree with maximum depth of 3. 
```{r}
library(rpart)
tree_model <- rpart(diabetes ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))
```
  
  - Calculate the accuracy of the model on the testing data. 
```{r}
pred <- predict(tree_model, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]
```
  
  - Plot the tree
```{r}
library(rattle)
fancyRpartPlot(tree_model)
```
  
  - Plot the variable importance by the tree
```{r}
barplot(tree_model$variable.importance)
```

-------

3. Practice Random Forest.  Do the follows: 

  - Use `randomForest` package, create a random forest of 1000 trees. 
```{r}
library(randomForest)
forest_model = randomForest(diabetes ~ ., data=df_train, ntree = 1000)
```
  
  - Calculate the accuracy of the model on the testing data. 
```{r}
pred <- predict(forest_model, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]
```
  
  - Plot the variable importance by the forest
```{r}
barplot(tree_model$variable.importance)
```

-------

4. Compare the accuracy of a forest of 1000 trees and a forest of 2000 trees. 
```{r}
forest_model = randomForest(diabetes ~ ., data=df_train, ntree = 1000)
pred <- predict(forest_model, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]

forest_model1 = randomForest(diabetes ~ ., data=df_train, ntree = 2000)
pred <- predict(forest_model1, df_test, type = "class")
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]
```

-------

5. Using Caret, create a tree with maximum depth of 3 and forest of 1000 trees. Compare the accuracy of these two models.
```{r}
library(caret)
model1 <- train(diabetes~., data=df_train, 
                method = "rpart2",
                maxdepth=3)
pred <- predict(model1, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]

model2 <- train(diabetes~., data=df_train, 
                method = "rf",
                ntree = 1000) 
pred <- predict(model2, df_test)
cm <- confusionMatrix(data = pred, reference = df_test$diabetes, positive = "pos")
cm$overall[1]
```

-------

6. Plot variable importance by the two models in 5. 
```{r}
plot(varImp(model1))
plot(varImp(model2))
```

