---
title: "Assignment9"
author: "Morgann Plante"
date: "10/12/2020"
output: html_document
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

-------

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint:

-  Use `filter` to select the most current date 

-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths
```{r}
library(tidyverse)
df <- read_csv('https://covid19.who.int/WHO-COVID-19-global-data.csv')
df %>% 
  filter(Date_reported== "2020-10-07") %>% 
  arrange(-Cumulative_deaths)
```


2. Make a plot revealing the number of deaths in the three countries with the most numbers of deaths
```{r}
library(gganimate)
library(ggplot2)
library(tidyverse)
library(knitr)
library(lubridate)
df %>% 
  filter(Country_code %in% c('US','BR','IN')) %>% 
  ggplot(aes(y=Cumulative_deaths,
            x=Date_reported,
            color=Country))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = Cumulative_deaths), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)
```

3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 
```{r}
df$death_per_cases <- df$Cumulative_deaths/df$Cumulative_cases
df %>% 
  filter(Date_reported== "2020-10-07") %>% 
  arrange(-death_per_cases)
```

4. Make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 
```{r}
df %>% 
  filter(Country_code %in% c('US','IT','MX')) %>% 
  ggplot(aes(y=death_per_cases,
            x=Date_reported,
            color=Country))+ 
  geom_line()+
  geom_point(size=3)+
  geom_text(aes(label = death_per_cases), 
            hjust = -.1, size=5) +
  transition_reveal(Date_reported)
```
