---
title: "Assignment6"
author: "Morgann Plante"
date: "9/28/2020"
output: html_document
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment6.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**Questions**

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. 
```{r}
library(tidyverse)
df <- read_csv('titanic.csv')

```

-------

1. For one continuous variable:

- Density
```{r}
df %>% ggplot()+
  geom_density(mapping = aes(x = Age))
#This plot shows the age of passengers because it is an important factor when evaluating who would be saved and who would not.
```

- Histogram 
```{r}
df %>% ggplot()+
  geom_histogram(mapping = aes(x = Age))
#This plot shows the age of passengers because it is an important factor when evaluating who would be saved and who would not.
```

- Boxplot
```{r}
df %>% ggplot()+
  geom_boxplot(mapping = aes(y = Age))
#This plot shows the age of passengers because it is an important factor when evaluating who would be saved and who would not.
```

-------

2. For one categorical variable

- Barplot
```{r}
df %>% ggplot()+
  geom_bar(mapping = aes(x = Sex))
#This plot shows the sex breakdown for the passengers because this is another important factor for evaluating who would be saved and who would not. In this time, it would be more likely for women and children to be saved over men.
```

-------

3. For one two continuous variables

- Scatter Plot
```{r}
df %>% ggplot()+
  geom_point(mapping = aes(x= Age, y= Fare))
#Age and far are evaluated for this plot because it shows that majority of the fares were under 100 making it more affoardable for the lower class.
```

- Line Plot
```{r}
df %>% ggplot()+
  geom_line(mapping = aes(x=Age, y=Fare))
#Age and far are evaluated for this plot because it shows that majority of the fares were under 100 making it more affoardable for the lower class.
```

- Smooth Plot
```{r}
df %>% ggplot()+
  geom_smooth(mapping = aes(x=Age, y=Fare))
#Age and far are evaluated for this plot because it shows the fares increased after arounf age 23 and started to decrease around age 52. This portion of ages is contained within the interquartile range of the first class, so it makes sense that the fares increased for this age range because the first class could afford it.
```

-------

4. For one continuous + one categorical variables

- Density
```{r}
df$Survived <- as.character(df$Survived)
df %>% ggplot()+
  geom_density(mapping = aes(x = Age , color = Survived))
#This plot shows the density of if someone survived or not based on ages. The plot is important because younger folks were more likely to be saved while the older passengers were not.
```

- BoxPlot
```{r}
df %>% ggplot()+
  geom_boxplot(mapping = aes(x = Age, y = Pclass))
#This plot is important because it shows the age breakdown for each class which helps the understanding of other important plots.
```

-------

5. For two categorical variables: barplot
```{r}
df$Pclass <- as.character(df$Pclass)
df %>% ggplot()+
  geom_bar(mapping=aes(x=Pclass, fill=Survived), position='dodge')
#This plot is important because it illustrates how the first class had more passengers survive than the third class even though the third class had more passengers in general.
```



