---
title: "Assignment7"
author: "Morgann Plante"
date: "9/30/2020"
output: html_document
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment7.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

**Questions**

Use the data of your own.  Produce the following types of plots and comment on each plot. Plots should be meaningful. If you use the data we used in class, make sure the plots are not the same as the ones in the slides. All plots should have title, caption, appropriate labels on x and y-axis. 
```{r}
library(tidyverse)
df <- read_csv('https://covidtracking.com/data/download/all-states-history.csv')
library(lubridate)
df$date <- ymd(df$date)
df$month = month(df$date)
df$weekday = wday(df$date)
df$monthday <- mday(df$date)
```

-------

1. Make a plot use `face_wrap` by one categorical variable (`face_wrap(~variable1)`)
```{r}
df %>% filter(state=='NY'|state=='RI') %>% 
  ggplot(mapping=aes(x=deathIncrease, y=hospitalizedIncrease))+ 
  geom_point()+
  geom_smooth()+
  facet_wrap(~state)+
  labs(x='Number of Deaths', 
       y = 'Number of Hospitalized', 
       title = 'Plot of Deaths and Hospitalizations', 
       caption = 'New York has a higher number of increased hospitalization and death in comparison to Rhode Island.')
```

2. Make a plot use `face_wrap` by two categorical variables (`face_wrap(variable1~variable2)`)
```{r}
df %>% filter(state=='NY'|state=='RI') %>% 
  ggplot(mapping=aes(x=positiveIncrease, y=hospitalizedIncrease))+ 
  geom_point()+
  geom_smooth()+
  facet_wrap(month~state)+
  labs(x='Number of Deaths', 
       y = 'Number of Hospitalized', 
       title = 'Plot of Deaths and Hospitalizations', 
       caption = 'The number of deaths and hospitalizations decrease for both states as the months progress.')
```

3. Make a plot use `position ='jitter'`. 
```{r}
df %>% filter(state=='RI'|state=="MA",deathIncrease>0) %>% ggplot()+ 
  geom_point(mapping=aes(x=deathIncrease, y=hospitalizedIncrease, color=state), position='jitter')+
  labs(x='Number of Deaths', 
       y = 'Number of Hospitalized', 
       color='State', 
       title = 'Scatter Plot of Deaths and Hospitalizations', 
       caption = 'Massachusetts has a higher rate of death per hospitalizations than Rhode Island.')
```

4. Make a plot use `position ='dodge'`.
```{r}
df %>% filter(month==9) %>% ggplot()+ 
  geom_bar(mapping=aes(x=weekday, fill=dataQualityGrade), 
           position = 'dodge')+
  labs(x='Day of the Week', 
       y = 'Number per Data Quality Grade', 
       color='Data Quality Grade', 
       title = 'Bar Chart of the Data Quality Grades per Weekday in September', 
       caption = 'The third day of the week has a higher count of "A" and "A+" grades than the other days of the week')
```

5. Make a plot use `position ='fill'`. 
```{r}
df %>% filter(month==9) %>% ggplot()+ 
  geom_bar(mapping=aes(x=weekday, fill=dataQualityGrade), 
           position = 'fill')+
  labs(x='Day of the Week', 
       y = 'Percentage of Data Quality Grade', 
       color='Data Quality Grade', 
       title = 'Bar Chart of the Data Quality Grades per Weekday in September', 
       caption = 'Each day of the week has approximately the same percentage for each data quatily grade.')
```

6. Make a plot use `geom_col`.
```{r}
df %>% group_by(month) %>% summarise(mean=mean(deathIncrease)) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=month, y=mean))+
  labs(x='Month', 
       y = 'Mean of Deaths', 
       title = 'Bar Chart of the Mean Number of Deaths for Each Month', 
       caption = 'There were more deaths on average in the fourth month.')
```

7. Change the theme of one of your plots. 
```{r}
df %>% filter(state=='RI'|state=="MA",deathIncrease>0) %>% ggplot()+ 
  geom_point(mapping=aes(x=deathIncrease, y=hospitalizedIncrease, color=state), position='jitter')+ theme_classic()+
    labs(x='Number of Deaths', 
       y = 'Number of Hospitalized', 
       color='State', 
       title = 'Scatter Plot of Deaths and Hospitalizations', 
       caption = 'Massachusetts has a higher rate of death per hospitalizations than Rhode Island.')
```

8. Save a plot and put a link of the plot to your Github Webpage.  To link a plot `abc.png` you just need to insert the following on the README.md

```{r}
df %>% filter(state=='RI'|state=="MA",deathIncrease>0) %>% ggplot()+ 
  geom_point(mapping=aes(x=deathIncrease, y=hospitalizedIncrease, color=state), position='jitter')+ theme_classic()+
    labs(x='Number of Deaths', 
       y = 'Number of Hospitalized', 
       color='State', 
       title = 'Scatter Plot of Deaths and Hospitalizations', 
       caption = 'Massachusetts has a higher rate of death per hospitalizations than Rhode Island.')+
  ggsave('abc.png')
```





