---
title: "Assignment 13"
author: "Morgann Plante"
date: "10/28/2020"
output: html_document
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](fa2020_assignment13.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Blackboard.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

#### NOTICE: Question 7 and 8 are optional. 

-------

1. Write a function to calculate the age of a person.  Give an example to test your function. 

- Input: a number, x, (year born)

- Output: Print out "You are r age". Where r is the age of the person, i.e. 2020 - x. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#29)
```{r}
age <- function(x)
{
  r = 2020 - x
  print(paste0('You are ', r, ' age'))
}
```
```{r}
age(1999)
```

-------

2. Write the following function. Give an example to test your function. 

- Input: a number

- Output: print out: "You input an even number!" if the number is event, or "You input an odd number!" otherwise. 

Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#33)
```{r}
numbers <- function(x)
{
  if(x %% 2 ==0)
  {
    print('You input an even number!') 
  }
  else
  {
    print('You input an odd number!')
  }
}
```
```{r}
numbers(39723)
```

-------

3. Write the following function.  Give an example to test your function. 

- Input: a numeric vector 

- Output:  

  - if the input vector has missing values: return the input vector with missing values replaced by mean
  
  - if the input vector has no missing value: return the same input vector
  
Hint: [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#37)
```{r}
mean_impute <- function(x)
{
  if(is.numeric(x))
  {
    mean_of_x <- mean(x, na.rm= TRUE)
    library(tidyr)
    x <- replace_na(x, mean_of_x) 
  }
return(x)    
}
```
```{r}
library(tidyverse)
df <- read_csv('titanic.csv')
x1 <- mean_impute(df$Age)
sum(is.na(x1))
```

-------

4. Write the following function.  Give an example to test your function. 

- Input: a vector x 

- Output:  The vector x where the missing values replaced by the mean (if x is numeric) or the mode (if x is non-numeric).  If x does not have missing value, return the same vector x.  

Hint: Combine the function in Question 3 and [this function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#37)
```{r}
mode_impute <- function(x)
{
  if(!is.numeric(x))
  {
    mode_of_x <- names(sort(-table(x)))[1]
    library(tidyr)
    x <- replace_na(x, mode_of_x) 
  }
  else
  {
    mean_of_x <- mean(x, na.rm= TRUE)
    library(tidyr)
    x <- replace_na(x, mean_of_x) 
  }
return(x)    
}
```
```{r}
library(tidyverse)
df <- read_csv('titanic.csv')
x1 <- mean_impute(df$Age)
x2 <- mode_impute(df$Embarked)
sum(is.na(x1))
sum(is.na(x2))
```

-------

5. Write the following function. Give examples to test your function. 

- Input: A data frame of two variables x and y

- Output: 

  - A boxplot of x by y if x is numeric and y is non-numeric
 
  - A boxplot of y by x if y is numeric and x is non-numeric
 
  - print out 'This function cannot visualize your data' otherwise

Hint: 

- [Similar Function](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#41)
 
- You can refer to this slide to plot a boxplot: https://bryantstats.github.io/math421/slides/6_viz.html#36
```{r}
viz <- function(d)
{
  if(is.numeric(d[[1]])& !(is.numeric(d[[2]])))
  {
    d %>% ggplot+ geom_boxplot(aes(x = d[[1]], y = d[[2]]))+
      labs(x = names(d)[1], y = names(d)[2])
  }
  else if (!(is.numeric(d[[1]]) & is.numeric(d[[2]])))
  {
    d %>% ggplot+ geom_boxplot(aes(x = d[[1]], fill = d[[2]]))+
      labs(x = names(d)[1], fill = names(d)[2])
  }
  else 
  {
    print('This function cannot visualize your data.')
  }
}
```
```{r}
d <- df %>% select(Age, Sex)
viz(d)
```
```{r}
d <- df %>% select(Age, Fare)
viz(d)
```
 
-------
 
6. Combine the function in Question 5 and the function in this example ([Link](https://bryantstats.github.io/math421/gganimate/13_writing_functions.html#41)) to have a function that can plot all possible cases (2 categorical, 2 numeric, 1 numeric and 1 categorical). 
```{r}
viz <- function(d)
{
  if(is.numeric(d[[1]])& !(is.numeric(d[[2]])))
  {
    d %>% ggplot+ geom_boxplot(aes(x = d[[1]], y = d[[2]]))+
      labs(x = names(d)[1], y = names(d)[2])
  }
  else if (!(is.numeric(d[[1]]) & is.numeric(d[[2]])))
  {
    d %>% ggplot+ geom_boxplot(aes(x = d[[1]], fill = d[[2]]))+
      labs(x = names(d)[1], fill = names(d)[2])
  }
  else if(is.numeric(d[[1]])&is.numeric(d[[2]]))
  {
    d %>% ggplot(aes(x = d[[1]], y = d[[2]]))+
      geom_point()+
      labs(x = names(d)[1], y = names(d)[2])
  }
  else if (!(is.numeric(d[[1]])|is.numeric(d[[2]])))
  {
    d %>% ggplot(aes(x = d[[1]], fill = d[[2]]))+
      geom_bar(position = 'dodge')+
      labs(x = names(d)[1], fill = names(d)[2])
  }
  else 
  {
    print('This function cannot visualize your data.')
  }
}
```
```{r}
d <- df %>% select(Age, Sex)
viz(d)
```
```{r}
d <- df %>% select(Age, Fare)
viz(d)
```

-------

#### NOTICE: Question 7 and 8 are optional. 

-------

7. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 

  - train_percent: a number presenting a proportion of training data. The default train_percent is .8

- Output: the accuracy of the decision model `rpart` where the training data is train_percent. 

-------

8. Write the following function. Give examples to test your function. 

- Input: 

  - input_data: a clean data frame with a variable name `target`. The `target` variable is also binary. 
```{r}
df = read_csv("titanic.csv")
names(df)[8] <- 'target'

df$PassengerId =  NULL
df$Ticket =  NULL
df$Name = NULL
df$Cabin = NULL

df$target <- factor(df$target)
df$Pclass = factor(df$Pclass)
df$Sex <- factor(df$Sex)
df$Embarked <- factor(df$Embarked)

df$Age[is.na(df$Age)] = mean(df$Age, na.rm = TRUE)
df = drop_na(df)
```

  - train_percent: a number presenting a proportion of training data. The default train_percent is .8
```{r}
modl2 <- function(input_data, train_percent = .8)
{
library(caret)
set.seed(00000)
splitIndex <- createDataPartition(input_data$target, p = train_percent, 
                                  list = FALSE)
df_train <- input_data[ splitIndex,]
df_test <- input_data[-splitIndex,]
library(rpart)
tree1<-rpart(target ~ ., 
             data = df_train)
library(rattle)
barplot(tree1$variable.importance)
}
```

- Output: the plot of variable important by random forest trained by caret. 
```{r}
modl2(df)
```



