---
title: "Midterm Presentation"
author: "Morgann Plante"
date: "10/26/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}

library(gapminder)
library(gganimate)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(knitr)
library(caret)
library(rpart)
df1 <- read_csv('midterm.csv')
df1$payfix <- NULL
df1$preopday<- NULL
df1$bwght <- NULL
df1$ecodub92 <- NULL
df1$obs_hour <- NULL
df1$nicu_day <- NULL
df1$diag_adm<- NULL
df1$pt_state<- NULL
df1$asource<- NULL
df1 <- drop_na(df1, raceethn, er_mode)
```


## Plot of Age of Patients Hospitalized

```{r}
library(tidyverse)
df1 %>% ggplot()+
  geom_histogram(mapping = aes(x = age))+
  labs(x='Age', 
       y = 'Count', 
       title = 'Age of Patients Hospitalized', 
       caption = 'There are more newborns than any other age hopitilized in Rhode Island in the given year.')
```
## Plot of the Sex of Patients Hospitalized
```{r}
df1$sex <- as.character(df1$sex)
df1 %>% ggplot()+
  geom_bar(mapping = aes(x = sex))+
  labs(x='Sex', 
       y = 'Count', 
       title = 'Sex of Patients Hospitalized', 
       caption = 'There are more females than males hopitilized in Rhode Island in this given year.')
```

## Plot of the Age and Total Patient Charges
```{r}
df1 %>%  ggplot()+
  geom_point(mapping = aes(x= age, y= tot), position="jitter")+
  labs(x='Age', 
       y = 'Total Patient Charhes', 
       title = 'Scatter Plot of Age and Total Patient Charges', 
       caption = 'Newborns have the largest patient charge.')
```

## Plot of the Ages of Patients with Race and Ethnicity
```{r}
df1$raceethn <- as.character(df1$raceethn)
df1 %>% ggplot()+
  geom_density(mapping = aes(x = age , color = raceethn))+
  labs(x='Age', 
       y = 'Count',
       color= 'Race and Ethnicity',
       title = 'Plot of Ages of Patients with Race and Ethnicity', 
       caption = 'White, or not Hispanic (1) patients are admitted to the hospital more after age 70 than other races and ethnicities.')
```

## Plot of Patients Race and Ethnicity and Sex
```{r}
df1 %>% ggplot()+
  geom_bar(mapping=aes(x=raceethn, fill=sex), position='dodge')+
  labs(x='Race and Ethnicity', 
       y = 'Count',
       color= 'Sex',
       title = 'Bar Plot of Patients Race and Ethnicity and Sex', 
       caption = 'White, or not Hispanic (1) female patients are admitted to the hospital more than other races and ethnicities.')
```

## Plot of Age and Total Patient Charges
```{r}
df1 %>% filter(raceethn==1|raceethn==7) %>% 
  ggplot(mapping=aes(x=age, y=tot))+ 
  geom_point()+
  geom_smooth()+
  facet_wrap(~raceethn)+
  labs(x='Age', 
       y = 'Total Patient Charges', 
       title = 'Plot of Age and Total Patient Charges', 
       caption = 'White, not Hispanics has higher total patient charges than Hispanics.')
```

## Plot of Age and Length of Stay
```{r}
df1 %>% 
  ggplot()+ 
  geom_point(mapping=aes(x=age, y=los, color=sex), position='jitter')+
  labs(x='Age', 
       y = 'Length of Stay', 
       color='Sex', 
       title = 'Scatter Plot Age and Length of Stay', 
       caption = 'Younger patients, less than 40, have more outliers for a longer length of stay than older patients.')
```

## Plot of the Number per Sex for the Month of Arrival
```{r}
df1 %>% ggplot()+ 
  geom_bar(mapping=aes(x=moa, fill=sex), 
           position = 'dodge')+
  labs(x='Month of Arrival', 
       y = 'Number per Sex', 
       color='Sex', 
       title = 'Bar Chart of the Number per Sex for the Month of Arrival', 
       caption = 'Females are admitted to the hospital more than males for any given month.')
```

## Plot of the Admission Type per Month of Admission
```{r}
df1$admtype <- as.character(df1$admtype)
df1 %>% ggplot()+ 
  geom_bar(mapping=aes(x=moa, fill=admtype), 
           position = 'fill')+
  labs(x='Month of Arrival', 
       y = 'Percentage of Admission Type', 
       color='Admission Type', 
       title = 'Bar Chart of the Admission Type per Month of Admission', 
       caption = 'The type of admission that is most common for any given month is Emergency, 1.')
```

## Plot of the Mean Total Patient Charges for Each Admission Type
```{r}
df1 %>% group_by(admtype) %>% summarise(mean=mean(tot)) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=admtype, y=mean))+
  labs(x='Admission Type', 
       y = 'Mean of Total Patient Charges', 
       title = 'Bar Chart of the Mean Total Patient Charges for Each Admission Type', 
       caption = 'Trauma, 5, has the greatest mean of total patient charges compared to other admission types.')
```

## Plot of the Mean Total Patient Charges for each Provider per Month of Admission
```{r}
d1 <- df1 %>% group_by(moa, provider) %>% summarise(mean = mean(tot))
d2 <- d1 %>% group_by(moa) %>% mutate(rank=rank(-mean)) %>% ungroup()
d3 <- d2 %>% filter(rank <= 7)
a1 <- d3 %>% ggplot(aes(x=rank, y=mean, group=provider, fill= provider, label=provider)) + geom_col()+
    geom_text(aes(y = mean, label = provider), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month of Arrival {closest_state}', x='', y='Total Patient Charges', fill='Provider', caption = 'Provider 7215, Bradley, has the most total patient charges on average than any provider for any given month.')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(moa)+
    ease_aes("cubic-in-out")
animate(a1, nframes = 400, fps = 10)
```
## Predictive Models
Target is the Total Patient Charges
- Low cost < 22050
- High cost > 22050
```{r, echo=FALSE}
median(df1$tot)
df1$target <- case_when(
  df1$tot <22050 ~ 'low cost',
  TRUE ~ 'high cost')

df2 <- select(df1,"age","sex","raceethn","provider","moa","mod","campus", "er_mode", "los","target")

library(caret)
df2$moa <- as.character(df2$moa)
df2$mod <- as.character(df2$mod)

set.seed(2020)
splitIndex <- createDataPartition(df2$target , p = .10, list = FALSE)
df_train <- df2[ splitIndex,]
df_test <- df2[-splitIndex,]

library(rpart)
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))
library(rattle)
fancyRpartPlot(tree_model)
barplot(tree_model$variable.importance)
```
## Tree Approach
```{r}
tuneGrid = expand.grid(maxdepth = 2:10)
trControl = trainControl(method = "cv",
                         number = 5)
tree_approach <- train(target~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach)
```
## Random Forest with Ranger Method
```{r}
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:5))
ranger_cv <- train(target~., data=df_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
plot(ranger_cv)
```
## Random Forest
```{r}
tuneGrid = expand.grid(mtry = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)
forest_cv <- train(target~., data=df_train, 
                                method = "rf", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(forest_cv)
```

## Comparison of the Models 
```{r}
results <- resamples(list(forest = forest_cv,
                          ranger = ranger_cv,
                          tree= tree_approach))
bwplot(results)
```

## Accuracy of the Final Model on the Test Data
```{r}
pred <- predict(ranger_cv, df_test)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test$target), positive = "high cost")
cm$overall[1]
```

## New Target Variable
The new target variable is the length of stay
- Short stay <3
- Long stay >3
```{r}
median(df1$los)
df1$target1 <- case_when(
  df1$los <3 ~ 'short stay',
  TRUE ~ 'long stay')

df3 <- select(df1,"age","sex","raceethn","provider","moa","mod","campus", "er_mode", "tot","target1")
```
```{r, echo=FALSE}
set.seed(2020)
splitIndex <- createDataPartition(df3$target1 , p = .10, list = FALSE)
df_train1 <- df3[ splitIndex,]
df_test1 <- df3[-splitIndex,]
```
## Variable Importance for New Target
```{r}
treeModel <- rpart(target1 ~ ., data = df_train1,
                 control = rpart.control(maxdepth = 3))
library(rattle)
fancyRpartPlot(treeModel)
barplot(treeModel$variable.importance)
```
## Tree Approach
```{r}
tuneGrid = expand.grid(maxdepth = 2:10)
trControl = trainControl(method = "cv",
                         number = 5)
treeApproach <- train(target1~., data=df_train1, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(treeApproach)
```
## Random Forest with Ranger Method
```{r}
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:5))
rangerCv <- train(target1~., data=df_train1, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
plot(rangerCv)
```
## Random Forest
```{r}
tuneGrid = expand.grid(mtry = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)
forestCv <- train(target1~., data=df_train1, 
                                method = "rf", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(forestCv)
```
## Comparison of the Models
```{r}
results <- resamples(list(forest = forestCv,
                          ranger = rangerCv,
                          tree= treeApproach))
bwplot(results)
```
## Accuracy of the Final Model on the Test Data
```{r}
pred <- predict(rangerCv, df_test1)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test1$target1), positive = "long stay")
cm$overall[1]
```