---
title: "Fall 2020 - Math 421 - Midterm"
author: "Morgann Plante"
date: "10/26/2020"
output: html_document
---

-------

# Instruction

The midterm has two components: the Rmarkdown notebook (html) and the presentation.  The notebook and the presentation file are due Nov 26. We will do the presentation in class 

**The notebook:** The notebook should be created using `rmarkdown` (like other assignments). The notebook should have a title. 


**The Presentation:** Present your results in 5-10 minutes. To make the presentation using Rmarkdown, do the follows: 

    - In Rstudio -> File -> New File -> R markdown
    
    - In the left panel, click to Presentation -> Click OK
    
    - Now you have an Rmarkdown that can be knitted to be a html presentation 

**What to present**:

  - 

**Data:**  The data for the mid-term project is the Rhode Island Department of Health Hospital Discharge Data.  Each row of the data presents a patient. 

Link: https://drive.google.com/open?id=15QNBf6YYKocK2nNIfpKDer58kQnCPNZJ 

-------

## I. Data Wranggling

1. Download the data file `hdd0318cy.sas7bdat`.  

2. Use `read_sas` in library `haven` to read the data. 
```{r}
library(haven)
df <- read_sas("hdd0318cy.sas7bdat")
```
    
3. Filter the data to have only patients of the year 2018 (`yod==2018`)
```{r}
library(tidyverse)
df <- filter(df, yod == 18)
```
    
4. Select to work with only following variables: 

```{r, eval=FALSE}
                      "yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day"
```
```{r}
library(tidyverse)
df1 <- select(df,"yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day")
write_csv(df1, 'midterm.csv')
```
 

*Notice*:  You may want to save the current data to your computer for easy access later.  To save the data file use `write_csv(df, 'midterm.csv')`, for example.  

5. What are variables that have missing values?
```{r}
library(tidyverse)
df1 <- read_csv('midterm.csv')
colSums(is.na(df1))
```

6. Remove all variables with missing values
```{r}
df1$payfix <- NULL
df1$preopday<- NULL
df1$bwght <- NULL
df1$ecodub92 <- NULL
df1$obs_hour <- NULL
df1$nicu_day <- NULL
df1$diag_adm<- NULL
df1$pt_state<- NULL
df1$asource<- NULL
df1 <- drop_na(df1, raceethn, er_mode)
```
```{r}
colSums(is.na(df1))
```

7. Refer to the data description in the file `HDD2015-18cy6-20-19.docx`, which variable recording the month of admission?, which variable recording the month of discharge?

The variable recording the month of admission is MOA. The variable recording the month of discharge is MOD.

8. Which month admitted the most number of patients? Which month admitted the most number of male patients?
```{r}
table(df1$moa)
#The month that admitted the most number of patients is May.

by(df1$sex==1, df1$moa, sum)
#The month that admitted the most number of male patients is October.
```

9. Which month has the most number of teenage female patients?
```{r}
library(tidyverse)
df1 %>% 
  filter(sex==2, age<18, age>12) %>% 
  group_by(moa) %>% 
  summarise(sum(moa))
#The month that has the most number of teenage female patients is December.
```

10. Which provider has the most number of female patients in October? 
```{r}
df2 <- filter(df1, moa==10)
by(df1$sex==2, df1$provider, sum)
#The provider that has the most number of female patients in October is provider 7205, Rhode Island Hospital.
```

11. Is female patients older than male patients, on average? 
```{r}
by(df1$age,df1$sex,mean)

#No, the female patients are not older than male patients on average.
```

12. Calculate the average age of patients by months. Which month has the oldest patients on average age?
```{r}
by(df1$age, df1$moa, mean)

#The month with the oldest patients on average is January.
```

13. What is the name of the provider that has the highest total charge?
```{r}
by(df1$tot, df1$provider, max)

#The provider with the highest total charge is provider 7214, Women and Infants.
```

14. What is the name of the provider that has the least total charge for teenage male on average?
```{r}
library(tidyverse)
df1 %>% 
  filter(sex==1, age<19, age>12) %>% 
  group_by(provider) %>% 
  summarise(mean(tot))

#The provider that has the least total charge for teenage males on average is provider 7206, Roger Williams.
```

15. Calculate the length of stays by races.  Which race has the longest length of stays on average?
```{r}
by(df1$los, df1$raceethn, mean, na.rm=TRUE)

#The race that has the longest length of stays on average is 9, unknown race and ethnicity.
```

16. On average, how much a 20 year-old male white get charged for staying 1 day?
```{r}
library(tidyverse)
df1 %>% 
  filter(sex==1, age==20, los==1, raceethn==1) %>% 
  summarise(mean(tot))
```

-------

## II. Data Visualization

Continue with the data from part I. 

1. Provides at least 10 meaningful plots. Comments on the plots. All plots should have title, caption, appropriate labels on x and y-axis
```{r}
library(tidyverse)
df1 %>% ggplot()+
  geom_histogram(mapping = aes(x = age))+
  labs(x='Age', 
       y = 'Count', 
       title = 'Age of Patients Hospitalized', 
       caption = 'There are more newborns than any other age hopitilized in Rhode Island in the given year.')
```
```{r}
df1$sex <- as.character(df1$sex)
df1 %>% ggplot()+
  geom_bar(mapping = aes(x = sex))+
  labs(x='Sex', 
       y = 'Count', 
       title = 'Sex of Patients Hospitalized', 
       caption = 'There are more females than males hopitilized in Rhode Island in this given year.')
```
```{r}
df1 %>%  ggplot()+
  geom_point(mapping = aes(x= age, y= tot), position="jitter")+
  labs(x='Age', 
       y = 'Total Patient Charhes', 
       title = 'Scatter Plot of Age and Total Patient Charges', 
       caption = 'Newborns have the largest patient charge.')
```
```{r}
df1$raceethn <- as.character(df1$raceethn)
df1 %>% ggplot()+
  geom_density(mapping = aes(x = age , color = raceethn))+
  labs(x='Age', 
       y = 'Count',
       color= 'Race and Ethnicity',
       title = 'Plot of Ages of Patients with Race and Ethnicity', 
       caption = 'White, or not Hispanic (1) patients are admitted to the hospital more after age 70 than other races and ethnicities.')
```
```{r}
df1 %>% ggplot()+
  geom_bar(mapping=aes(x=raceethn, fill=sex), position='dodge')+
  labs(x='Race and Ethnicity', 
       y = 'Count',
       color= 'Sex',
       title = 'Bar Plot of Patients Race and Ethnicity and Sex', 
       caption = 'White, or not Hispanic (1) female patients are admitted to the hospital more than other races and ethnicities.')
```
```{r}
df1 %>% filter(raceethn==1|raceethn==7) %>% 
  ggplot(mapping=aes(x=age, y=tot))+ 
  geom_point()+
  geom_smooth()+
  facet_wrap(~raceethn)+
  labs(x='Age', 
       y = 'Total Patient Charges', 
       title = 'Plot of Age and Total Patient Charges', 
       caption = 'White, not Hispanics has higher total patient charges than Hispanics.')
```
```{r}
df1 %>% 
  ggplot()+ 
  geom_point(mapping=aes(x=age, y=los, color=sex), position='jitter')+
  labs(x='Age', 
       y = 'Length of Stay', 
       color='Sex', 
       title = 'Scatter Plot Age and Length of Stay', 
       caption = 'Younger patients, less than 40, have more outliers for a longer length of stay than older patients.')
```
```{r}
df1 %>% ggplot()+ 
  geom_bar(mapping=aes(x=moa, fill=sex), 
           position = 'dodge')+
  labs(x='Month of Arrival', 
       y = 'Number per Sex', 
       color='Sex', 
       title = 'Bar Chart of the Number per Sex for the Month of Arrival', 
       caption = 'Females are admitted to the hospital more than males for any given month.')
```
```{r}
df1$admtype <- as.character(df1$admtype)
df1 %>% ggplot()+ 
  geom_bar(mapping=aes(x=moa, fill=admtype), 
           position = 'fill')+
  labs(x='Month of Arrival', 
       y = 'Percentage of Admission Type', 
       color='Admission Type', 
       title = 'Bar Chart of the Admission Type per Month of Admission', 
       caption = 'The type of admission that is most common for any given month is Emergency, 1.')
```
```{r}
df1 %>% group_by(admtype) %>% summarise(mean=mean(tot)) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x=admtype, y=mean))+
  labs(x='Admission Type', 
       y = 'Mean of Total Patient Charges', 
       title = 'Bar Chart of the Mean Total Patient Charges for Each Admission Type', 
       caption = 'Trauma, 5, has the greatest mean of total patient charges compared to other admission types.')
```

2. Make an animation
```{r}
library(gganimate)
library(ggplot2)
library(tidyverse)
library(knitr)
library(lubridate)

d1 <- df1 %>% group_by(moa, provider) %>% summarise(mean = mean(tot))
d2 <- d1 %>% group_by(moa) %>% mutate(rank=rank(-mean)) %>% ungroup()
d3 <- d2 %>% filter(rank <= 7)
a1 <- d3 %>% ggplot(aes(x=rank, y=mean, group=provider, fill= provider, label=provider)) + geom_col()+
    geom_text(aes(y = mean, label = provider), hjust = 1.4)+ 
    coord_flip(clip = "off", expand = FALSE) +scale_x_reverse()+
    labs(title = 'Month of Arrival {closest_state}', x='', y='Total Patient Charges', fill='Provider', caption = 'Provider 7215, Bradley, has the most total patient charges on average than any provider for any given month.')+
    theme(plot.title = element_text(hjust = 1, size = 22),
          axis.ticks.y = element_blank(),
          axis.text.y  = element_blank()) + 
    transition_states(moa)+
    ease_aes("cubic-in-out")
animate(a1, nframes = 400, fps = 10)
```

-------

## III. Predictive Models

Continue with the data from part I. Use the follows as the target and input variables: 

*Target Variable*: Create the target variable taking value of 

  - `low cost` if the total charge of a patient (`tot`) is smaller than the median of the total charge, and

  - `high cost` otherwise. 
```{r}
median(df1$tot)
df1$target <- case_when(
  df1$tot <22050 ~ 'low cost',
  TRUE ~ 'high cost')
```


*Input Variables*:

  - "age","sex","raceethn","provider","moa","mod","campus", "er_mode", 'los'
```{r}
df2 <- select(df1,"age","sex","raceethn","provider","moa","mod","campus", "er_mode", "los","target")
```

-------

1. Make sure all the categorical variables are factor, numeric variables are numeric. Set Training : Testing Split = 10 : 90 
```{r}
library(caret)
df2$moa <- as.character(df2$moa)
df2$mod <- as.character(df2$mod)

set.seed(2020)
splitIndex <- createDataPartition(df2$target , p = .10, list = FALSE)
df_train <- df2[ splitIndex,]
df_test <- df2[-splitIndex,]
```

2. Train a decision tree using `rpart`.  Plot the decision tree. Plot the variable importance ranked by the tree. 
```{r}
library(rpart)
tree_model <- rpart(target ~ ., data = df_train,
                 control = rpart.control(maxdepth = 3))
library(rattle)
fancyRpartPlot(tree_model)
barplot(tree_model$variable.importance)
```

3. Using caret for this question. Set `Training Control` to be: Use Cross-Validation of 5 folds across all models.  Train & tune at least 3 different models (i.e. three different values for `method=` in the train function of caret).  Plot the hyper-parameter tuning plots for each model. 
```{r}
library(caret)

tuneGrid = expand.grid(maxdepth = 2:10)
trControl = trainControl(method = "cv",
                         number = 5)
tree_approach <- train(target~., data=df_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(tree_approach)
```
```{r}
library(rpart)
library(rpart.plot)
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:5))
ranger_cv <- train(target~., data=df_train, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
plot(ranger_cv)
```
```{r}
tuneGrid = expand.grid(mtry = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)
forest_cv <- train(target~., data=df_train, 
                                method = "rf", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(forest_cv)
```

4. Plot the comparison of the models in 2. 
```{r}
results <- resamples(list(forest = forest_cv,
                          ranger = ranger_cv,
                          tree= tree_approach))
bwplot(results)
```

5. What is your final selection for the model? Test the accuracy of your final model on the test data. 
```{r}
pred <- predict(ranger_cv, df_test)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test$target), positive = "high cost")
cm$overall[1]
```

6. Create another `target` variable (binary), decide the input variables and redo 1 to 5. 
```{r}
median(df1$los)
df1$target1 <- case_when(
  df1$los <3 ~ 'short stay',
  TRUE ~ 'long stay')

df3 <- select(df1,"age","sex","raceethn","provider","moa","mod","campus", "er_mode", "tot","target1")
```
```{r}
set.seed(2020)
splitIndex <- createDataPartition(df3$target1 , p = .10, list = FALSE)
df_train1 <- df3[ splitIndex,]
df_test1 <- df3[-splitIndex,]
```
```{r}
library(rpart)
treeModel <- rpart(target1 ~ ., data = df_train1,
                 control = rpart.control(maxdepth = 3))
library(rattle)
fancyRpartPlot(treeModel)
barplot(treeModel$variable.importance)
```
```{r}
library(caret)

tuneGrid = expand.grid(maxdepth = 2:10)
trControl = trainControl(method = "cv",
                         number = 5)
treeApproach <- train(target1~., data=df_train1, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(treeApproach)
```
```{r}
library(tidyverse)
library(caret)
library(rpart)
trControl = trainControl(method = "cv",
                         number = 5)
tuneGrid = expand.grid(mtry = 2:4,
                       splitrule = c('gini', 'extratrees'),
                       min.node.size = c(1:5))
rangerCv <- train(target1~., data=df_train1, 
                    method = "ranger", 
                    trControl = trControl,
                    tuneGrid = tuneGrid)
plot(rangerCv)
```
```{r}
tuneGrid = expand.grid(mtry = 2:5)
trControl = trainControl(method = "cv",
                         number = 5)
forestCv <- train(target1~., data=df_train1, 
                                method = "rf", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)
plot(forestCv)
```
```{r}
results <- resamples(list(forest = forestCv,
                          ranger = rangerCv,
                          tree= treeApproach))
bwplot(results)
```
```{r}
pred <- predict(rangerCv, df_test1)
cm <- confusionMatrix(data = factor(pred), reference = factor(df_test1$target1), positive = "long stay")
cm$overall[1]
```

-------